---
title: "Investigating COVID-19 Virus Trends"
author: "ngclnh"
output: html_notebook
---
In this project, we will analyze the data from covid19.csv to answer the question: "Which countries have had the highest number of positive cases against the number of tests?"

To do that, first, we need to explore our data

### Exloring the data

```{r}
# Read the csv file
library(readr)

covid <- read_csv("covid19.csv")
```


```{r}
# The dimension of the data
dim(covid)
```


```{r}
# Column names
cols <- colnames(covid)
cols
```


```{r}
# Show the first few rows
head(covid)
```
```{r}
# Show the summary of the dataset
library(tibble)

glimpse(covid)
```

Now that we get a glimpse of our dataset, we will extract the data that relevant to answer the question.

### Isolating that data we need

Notice that for the Province_State column, there's a mix of country-level and state-level. We will now extract only "All States" data

```{r}
library(dplyr)

# Filter the "All States" Province states and remove the `Province_State` column
covid_all_states <- covid %>% 
  dplyr::filter(Province_State == "All States") %>% 
  select(-Province_State) # Remove the Province_State column as now we know our new dataset only contains 'All State' for Province_State column

head(covid_all_states)
```
Let's recall the description of the dataset's columns.

1. `Date`: Date
2. `Continent_Name`: Continent names
3. `Two_Letter_Country_Code`: Country codes
4. `Country_Region`: Country names
5. `Province_State`: States/province names; value is `All States` when state/provincial level data is not available
6. `positive`: Cumulative number of positive cases reported.
7. `active`: Number of actively cases on that **day**.
8. `hospitalized`: Cumulative number of hospitalized cases reported.
9. `hospitalizedCurr`: Number of actively hospitalized cases on that **day**.
10. `recovered`: Cumulative number of recovered cases reported.
11. `death`: Cumulative number of deaths reported.
12. `total_tested`: Cumulative number of tests conducted.
13. `daily_tested`: Number of tests conducted on the **day**; if daily data is unavailable, daily tested is averaged across number of days in between.
14. `daily_positive`: Number of positive cases reported on the **day**; if daily data is unavailable, daily positive is averaged across number of days in.

Notice that there's cumulative and daily data. We should work with those data separately. 

We will choose to work with daily data

```{r}
# Select only Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive columns
covid_all_states_daily <- covid_all_states %>% 
  select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)

head(covid_all_states_daily)
```
# Analyze the data

We want to find the top ten countries based on tested cases

```{r}
# Summarizing the data based on the `Country_Region` column.
covid_all_states_daily_sum <- covid_all_states_daily %>% 
  group_by(Country_Region) %>% 
  summarise(tested = sum(daily_tested), 
            positive = sum(daily_positive),
            active = sum(active),
            hospitalized = sum(hospitalizedCurr)) %>% 
  arrange(desc(tested)) #arrange the tested column in descending order

covid_all_states_daily_sum
```
```{r}
covid_top_10 <- head(covid_all_states_daily_sum, 10)

covid_top_10
```

Let's now dentify the Highest Positive Against Tested Cases

```{r}
# Getting vectors
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested
positive_cases <- covid_top_10$positive
active_cases <- covid_top_10$active
hospitalized_cases <- covid_top_10$hospitalized

# Naming vectors with the country names
names(positive_cases) <- countries
names(tested_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
```

```{r}
positive_cases
sum(positive_cases)
mean(positive_cases)
positive_cases/sum(positive_cases)
```

```{r}
positive_cases/tested_cases
```

```{r}
# The top three positive against tested cases
positive_tested_top_3 <- c("United Kingdom" = 0.11, "United States" = 0.10, "Turkey" = 0.08)
```

# Keeping relevant information

```{r}
# Creating vectors
united_kingdom <- c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0, 0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)

# Creating the matrix covid_mat
covid_mat <- rbind(united_kingdom, united_states, turkey)

# Naming columns
colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")

#Displaying the matrix
covid_mat
```

```{r}
question <- "Which countries have had the highest number of positive cases against the number of tests?"

answer <- c("Positive tested cases" = positive_tested_top_3)

datasets <- list(
  original = covid,
  allstates = covid_all_states,
  daily = covid_all_states_daily,
  top_10 = covid_top_10
)

matrices <- list(covid_mat)
vectors <- list(cols, countries)

data_structure_list <- list("dataframe" = datasets, "matrix" = matrices, "vector" = vectors)

covid_analysis_list <- list(question, answer, data_structure_list)

covid_analysis_list[[2]]
```
```{r}
question
answer
```

